Continuous developmental change explains discontinuities in word learning

### 2019-08-27 journal/ms.Rmd
first version journal

 "Cognitive development is often characterized in term of discontinuities, but these discontinuities can sometimes be apparent rather than actual and can arise from continuous developmental change. To explore this idea, we use as a case study the finding by Stager and Werker (1997) that children's early ability to distinguish similar sounds does not automatically translate into word learning skills. Early explanations proposed that children may not be able to encode subtle phonetic contrasts when learning novel word meanings, thus suggesting a discontinuous/stage-like pattern of development. However, later work has revealed (e.g., through using more precise testing methods) that children do encode such contrasts, thus favoring a continuous pattern of development. Here we propose a probabilistic model describing how development may proceed in a continuous fashion across the lifespan. It characterizes developmental change as improvement in the precision of perceptual encoding. Our model explained previous findings in the literature and provided a new prediction --- the referents' visual similarity modulates word leanring accuracy. The models' predictions were corroborated by human data we collected from both preschool children and adults. The braoder impact of this work is to show that computational models, such as ours, can help us explore the extent to which episodes of cogntive development that are typically thought of as discontinuities may emerge from simpler, continuous mechanisms."

@@

Nevertheless, at least in some cases, development may only appear to be stage-like. Some discontinuities may be related to how we measure a specific skill --- some tasks may not allow us to detect successful learning.  Other discontinuities may emerge due to statistical thresholding (e.g., p-value < 0.05) which can create a spurious dichotomy between success and failure in observing a given behavior. In such cases, positing discontinuous stages is unnecessary. Instead, a continuous model --- involving similar representations across the lifespan --- may provide a simpler and more transparent account of development.

We use a case study from word learning literature. @stager1997 first showed that children's early ability to distinguish similar sounds does not automatically translate into word learning skills. The authors measured word learning using an audio-visual habituation Switch task. First, infants are familiarized with two word-object pairings (e.g., lable A with object A and label B with object B). Second, they are tested using two types of trials. The control 'same' trial consists of a correct pairing  (e.g., label A with object A) and the 'switch' trial consists of a wrong pairing (e.g., label A with object B). If babies have correctly learned the association during the familiarization, they are supposed to be surprised by the 'switch' trial and not by the 'same' trial. The former should thus result in a greater looking time compared to the latter [@werker1998].

@@

The initial discontinuous interpretation has been challenged by subsequent work. For instance, @yoshida2009 investigated whether failure in the switch task reflects a lack of sound encoding during *familiarization*, or whether it is only due to the demands of the *testing* method which does not allow learning below a certain threshold to be detected. They used the same familiarization procedure as @stager1997, but instead of comparing the looking times in 'same' and 'switch' trials, they tested infants using a two-alternative choice task comparing fixations to target and distractor objects [@golinkoff1987; @fernald2006]. Using this easier and more precise task, the researcher found evidence for learning even in 14-month olds.

@@

Some researchers [@swingley2007; @yoshida2009; @pajak2016] proposed that word knowledge may not be encoded in a binary fashion, i.e., it is not the case that children either succeed or fail in encoding minimal contrast when learning the meanings. Rather, they may be encoding this knowledge in a graded fashion (ses also munakata, 2001). According to this view, development does not so much involve a qualitative shift (i.e., a sudden emergence of an ability that did not exist before) as much as it consists in the continuous refinement of initially noisy knowledge.

@@

This general framework --- in which the precision of word knowledge is characterize with the variance of a probability distribution --- can already provide an intuitive way of thinking about several findings. In particular, unlike the binary view, the probabilistic view allows for the possibility that knowldege can be successful while still being noisy. This new understanding can provide an account for the fact that children show evidence of learning in some testing condition [@yoshida2009], but not in others [@stager1997].

In a word-pair learning paradigm, children are supposed to associate one label, e.g, "bih", with object 1 and a second label, e.g., "dih", with object 2.  Infants may succeed in learning both associations. Nevertheless, the precision with which the words are encoded can still be large, causing their probability distributions to overalp (Figure \@ref(fig:illus), top). The way this knowledge is tested can lead to divering results.

Failure in the Switch task [@stager1997] can be exlained as follows. Children are understood to succeed if they reject the wrong pairing (e.g., "bih" with object 2). However, the overlap means that children could think of "bih" as plausible misprononciation of "dih". They may require a stronger sound mismatch to confidently reject the possibility that the wrong label is just an acceptable deviation from the tyical pronouciation. In contrast, success in the two-alternative choice task [@yoshida2009] occurs precisely becasue children do not have to reject a wrong pairing. Instead, the setting allows them to just choose which objects is a better match to a given label, thus revealing whether they have a prefernce, albeit samll, for the correct pairing.

In addition to explaining behavior in different tasks, the probabilsitic understanding also accounts for different behaviors in the same task. In particular, when the labels are quite distinct in the pereptual space ("lif" vs. "neem"), the distributions do not overlap as much as in the case of similar-sounding words (Figure \@ref(fig:illus), left). This fact means that the learners will have a stronger preference fot the correct label-object pairings and less tolerance for the wrong ones. That said, perceptual distinctiveness can be enhanced even for minimally different sounds when other cues highlight their difference [@rost2009; @rost2010; @thiessen2007; @yeung09; @dautriche2015].

Finally, development can be understood as an increase in the precision (i.e., a decrease in the variance) of the probabilistic representations, leading to a lower overlap between the similar-sounding words (Figure \@ref(fig:illus), right).  Importantly, a more precise representation still has a non-zero variance --- Learning difficulties can still be induced with challenging stimuli or in cognitively demanding situations as was demonstrate in adults studies [@white2013; @pajak2016].

@@

The present work proposes a model of word-pair learning based on the probabilistic account. We tested the ability of this model to both *explain* various findings in previous experiments in both children and adults (e.g., the fact that similar words are harder to learn than different words) and to *predict* new learning patterns that have not been tested before (i.e., the effect of the referents' similarity on word learning). The broader impact of this work is to explore the extent to which development can be understood as a continuous refinement in similar representations across the lifespan.

The paper is organized as follows. First, we introduce the model and we explain how it allows us to characterize behavior in a word learning task which resembles the one used in @stager1997 and @yoshida2009. Then we explore the predictions of the model through simulating its behavior across different parameter settings. Next we quantify the extent to which the model's predictions account for human data we collected from both preschool children and adults. Finally, we discuss the results in the lights of existing accounts of word development.

```{r illus, fig.cap = "An illustration of the probabilistic/continuous account using simulated data. A word is represented with a distribution over the perceptual space (indicated in red or blue). When the uncertainty of the representation is large relative to the distance between the stimuli (top panel), an instance of the red category (indicated with a star) could also be a plausible instance of the green category, hence the low recognition accuracy score. The accuracy increases when the stimuli are less similar (left panel), or when the representation are more precise (right panel).", fig.align = "center", out.width = "300px"}
knitr::include_graphics("figs/illustration.png")
```

@@

```{r model, fig.cap = "Graphical representation of our model. Circles indicate random variables (shading indicates observed variables). The squares indicate fixed model parameters.", fig.align = "center", out.width = "300px"}
knitr::include_graphics("figs/model.png")
```

@@

$$ p(o | C) \sim  \mathcal{N}(\mu_C, \sigma^2_C) $$

$$ p(s| L) \sim  \mathcal{N}(\mu_L, \sigma^2_L) $$

Finally, we assume there to be one-to-one mappings between concepts and labels and that observers have successfully learned these mappings during the exposure phase:
$$
P(L_i|C_j) = 
\begin{cases}
  1 & \text{if  }  i=j \\  
  0  & \text{otherwise  }
\end{cases}
$$

@@

The learner hears a sound $s$ and has to decide which object $o$ provides an optimal match to this sound (see Figure \@ref(fig:task)). To this end, they must compute the probability $P(o|s)$ for all possible objects. This probability can be computed by summing over all possible concepts and labels:
$$P(o|s)=\sum_{C,L} P(o, C, L| s) \propto \sum_{C,L} P(o, C, L, s) $$

The joint probability $P(o, C, L, s)$ is obtained by factoring the Bayesian network in Figure \@ref(fig:model):
$$P(o,C,L,s) = P(s|L)P(L|C)P(C|o)P(o) $$

which can be transformed using Bayes rule into:

$$P(o,C,L,s) = P(s|L)P(L|C)P(o|C)P(C) $$

Finally, assuming that the concepts' prior probability is uniformly distributed\footnote{This is a reasonable assumption in our particular case given the similarity of the concepts used in each naming situation in our experiment.}, we obtain the following expression, where all conditional dependencies are now well defined:

$$
P(o|s) = \frac{\sum_{C,L} P(s|L)P(o|C)P(L|C)}{\sum_{o} \sum_{C,L} P(s|L)P(o|C)P(L|C)}
$$

@@

We use the model to predict word learning in a task similar to the one introduced by @stager1997. We used a modified version of the task where the testing method consists in a two-alternative forced choice [@yoshida2009]. In this task, participants are first exposed to the association between pairs of nonsense words (e.g., "lif"/"neem") and pairs of objects. The word-object associations are introduced sequentially. After this exposure phase, participants perform a series of test trials. In each of these trials, one of the two sounds is uttered (e.g., “lif”) and participants choose the corresponding object from the two alternatives. An overview of the task is shown in Figure \@ref(fig:task).

From the general expression (1), we derive three exact analytical solutions instantiating different learning assumptions. The first solution is derived by assuming that the labels are recovered from sounds with a certain level of uncertainty $\sigma_L$, but that concepts are unambiguously recovered from the observed objects, i.e., $\sigma_C \rightarrow 0$. This assumption has been made --- whether implicitly or explicitly --- by most previous work in this line of research. One important implication of this assumption is that only the similarity of word sounds modulates success in word learning, not the similarity of the referents (as long as these referents are differentiated perceptually). This assumption yields the following probability function:

\begin{equation}
P(o_T|s)= \frac{1}{1 + e^{-\frac{\Delta s^2}{2\sigma_L^2}}}
\end{equation}

The second solution is derived by making the more general assumption that both the labels and the concepts are recovered with ambiguity from the sounds and objects. We first introduce the simplifying assumption that the label-related uncertainty $\sigma_L$ and the concept-related uncertainty $\sigma_C$ are of a similar magnitude, i.e., $\sigma_C \approx \sigma_L = \sigma$. This assumption makes the prediction that the sound similarity and the object similarity impact word learning accuracy in exactly the same way. Furthermore, it allows us to study the behavior of the model with minimal free parameters.

\begin{equation}
P(o_T|s)= \frac{1 + e^{- \frac{\Delta s^2 + \Delta o^2}{2\sigma^2}}}{1 + e^{-\frac{\Delta s^2 + \Delta o^2}{2\sigma^2}}+ e^{-\frac{\Delta s^2}{2\sigma^2}} + e^{-\frac{\Delta o^2}{2\sigma^2}}}
\end{equation}

We finally derive the third (and most general) solution which allows label- and concept-related uncertainties to vary independently.

\begin{equation}
P(o_T|s)= \frac{1 + e^{- (\frac{\Delta s^2}{2\sigma_L^2}+ \frac{\Delta o^2}{2\sigma_C^2})}}{1 + e^{-(\frac{\Delta s^2}{2\sigma_L^2}+ \frac{\Delta o^2}{2\sigma_C^2})}+ e^{-\frac{\Delta s^2}{2\sigma_L^2}} + e^{-\frac{\Delta o^2}{2\sigma_C^2}}}
\end{equation}

In order to understand the predictions of the models (especially the more general ones, i.e., Model 2 and 3), Figure \ref{fig:simulation} show some simulations of the accuracy $P(o_T|s)$ as a function of the distinctiveness parameters ($\Delta s$ and $\Delta o$) and the uncertainty parameters $\sigma_L$ and $\sigma_C$.

@@

```{r simulation, fig.cap = "The predicted probability of accurate responses in the testing phase as a function of stimuli distinctiveness $\\Delta s$ and $\\Delta o$ and representation precision $\\sigma$ (For clarity, we assume here that $\\sigma$=$\\sigma_C$=$\\sigma_L$). Dashed line represents chance."}
plot_simulation

@@

2) For fixed values of $\Delta s$ and $\Delta o$, accuracy increases when the representational uncertainty $\sigma$ decreases. This fact may explain development, i.e., younger children have noisier representations [see @swingley2007; @yoshida2009], which leads to lower word recognition accuracy, especially for similar-sounding words.

@@

### Participants
We report data from $N=$ `r N_good_ch` children ages 4-5 years from the Bing Nursery School on Stanford University’s campus. An additional $N=$ `r N_all_ch-N_good_ch` children participated but were removed from analyses because they were not above chance on the catch trials due to the challenging nature of our procedure (see below). We also report data from $N=$ `r N_good_ad` adult participants tested on Amazon Mechanical Turk. An additional $N=$ `r N_all_ad-N_good_ad` were tested but removed from analyses because they had low scores on the catch trials or because they were familiar with the non-English sound stimuli we used in the adult experiment.

@@

As for the objects, we used the Dynamic Stimuli javascript library\footnote{https://github.com/erindb/stimuli} which allowed us to generate objects in four different categories: "tree," "bird," "bug," and "fish." These categories are supposed to be naturally occurring kinds that might be seen on an alien planet. In each category, we generated *different*, *intermediate* and *similar* pairs by manipulating a continuous property controlling features of the category's shape (e.g, body stretch or head fatness).

In order to validate and quantify our similarity scales, we ran a separate survey on Amazon Mechanical Turk where we asked $N=20$ adults participants to evaluate the similarity of each sound and object pair on a 7-point scale. Data are shown in Figure \@ref(fig:stim) where we scaled responses within the range [0,1] for each stimulus group. These data will be used in all models as a proxy for the perceptual distance between the sound pairs ($\Delta s$) and the object pairs ($\Delta o$).

@@

sim_plot <- ggplot(sim_transform, 
      aes(x = labels, y = mean, group = factor(stimuli), col=factor(stimuli))) +
  #geom_point(size=2)+
  geom_line()+
  geom_pointrange(aes(ymin = lower, ymax = upper), 
                 position = position_dodge(width = .1),
                 size=0.2) + 
  scale_colour_discrete(name="Stimuli")+
  theme_few()+
theme(legend.title = element_text(size=7.5),
      legend.text=element_text(size=7.5),
      axis.text = element_text(size = 7.5),
      strip.text = element_text(size = 7.5),
      axis.title = element_text(size = 7.5),
      aspect.ratio = 0.7
      ) +
  ylim(c(0,1.1)) +
  xlab("Stimuli group")+ ylab("Perceptual distance (experimental)")

@@

Each age group saw only two of the three levels of similarity described in the previous sub-section: *different* vs. *intermediate* for the preschoolers, and *intermediate* vs. *similar* for adults. We made this choice in light of pilot studies showing that adults were at ceiling with *different* sounds/objects, and children were at chance with the *similar* sounds/objects. That said, this difference in the level of similarity is accounted for in the model through using the appropriate distance (Figure \@ref(fig:stim)).

```{r stim, fig.env = "figure", fig.pos = "h", fig.align='center', fig.width=3.5, fig.height=2.5, fig.cap = "Distances for both sound and object pairs from an adult norming study. Data represent Likert values normalized to [0,1] interval. Error bars represent 95\\% confidence intervals."}

@@

To maximize our ability to measure subtle stimulus effects, the experiment was a 2x2 within-subjects factorial design with four conditions: high/low sound similarity crossed with high/low visual object similarity. Besides the four conditions, we also tested participants on a fifth catch condition which was similar in its structure to the other ones but was used only to select participants who were able to follow the instructions and show minimal learning.

@@

Preschoolers were tested at the nursery school using a tablet, whereas adults used their own computers to complete the same experiment online. Participants were tested in a sequence of five conditions: the four experimental conditions plus the catch condition. In each condition, participants saw a first block of four exposure trials followed by four testing trials, and a second block of two exposure trials (for memory refreshment) followed by an additional four testing trials. The length of this procedure was demanding, especially for children, but we adopted a fully within-subjects design based on pilot testing that indicated that precision of measurement was critical for testing our experimental predictions.

In the exposure trials, participants saw two objects associated with their corresponding sounds. We presented the first object on the left side of the tablet's screen simultaneously with the corresponding sound. The second sound-object association followed on the other side of the screen after 500ms. For both objects, visual stimuli were present for the duration of the sound clip (about 800ms).  In the testing trials, participants saw both objects simultaneously and heard only one sound. They completed the trial by selecting which of the two objects corresponded to the sound. The object-sound pairings were randomized across participants, as was the order of the conditions (except for the catch condition which was always placed in the middle of the testing sequence). We also randomized the on-screen position (left vs. right) of the two pictures on each testing trial.

@@

```{r allData, fig.env = "figure", fig.pos = "h", fig.align='center', fig.cap = "Accuracy of word recognition as a function of the sound distance, the object distance, and the age group (preschool children vs. adults). We show both the models' predictions (dashed lines) and the experimental results (solid lines). The single-variance model uses one joint fitting parameter for both sound and meaning variances. The double-variance model uses two separate fittings parameters for the sound and the meaning variances. Error bars represent 95\\% confidence intervals."}

>>>>

 "Cognitive development is often characterized in terms of discontinuities, but these discontinuities can sometimes be apparent rather than actual and can arise from continuous developmental change. To explore this idea, we use as a case study the finding by Stager and Werker (1997) that children's early ability to distinguish similar sounds does not automatically translate into word learning skills. Early explanations proposed that children may not be able to encode subtle phonetic contrasts when learning novel word meanings, thus suggesting a discontinuous/stage-like pattern of development. However, later work has revealed (e.g., through using more precise testing methods) that children do encode such contrasts, thus favoring a continuous pattern of development. Here we propose a probabilistic model that represents word knowledge in a graded fashion and characterizes developmental change as improvement in the precision of this graded knowledge. Our model explained previous findings in the literature and provided a new prediction --- the referents' visual similarity modulates word learning accuracy. The models' predictions were corroborated by human data we collected from both preschool children and adults. The broader impact of this work is to show that computational models, such as ours, can help us explore the extent to which episodes of cognitive development that are typically thought of as discontinuities may emerge from simpler, continuous mechanisms."

@@

Nevertheless, at least in some cases, development may only appear to be stage-like. Some discontinuities may be related to how we measure a specific skill. Other discontinuities may emerge due to statistical thresholding (e.g., p-value < 0.05) which can create a spurious dichotomy between success and failure in observing a given behavior. In such cases, positing discontinuous stages is unnecessary. Instead, a continuous model --- involving similar representations across the lifespan --- may provide a simpler and more transparent account of development (see also @McMurray2007 and @shultz1995).

We use a case study from word learning literature. @stager1997 first showed that children's early ability to distinguish similar sounds does not automatically translate into word learning skills. The authors measured word learning using an audio-visual habituation Switch task. First, infants are familiarized with two word-object pairings (e.g., label 1 with object 1 and label 2 with object 2). Second, they are tested using two types of trials. The control 'same' trial consists of a correct pairing  (e.g., label 1 with object 1) and the 'switch' trial consists of a wrong pairing (e.g., label 1 with object 2). If babies have correctly learned the association during the familiarization, they are supposed to be surprised by the 'switch' trial and not by the 'same' trial. The former should thus result in a greater looking time compared to the latter [@werker1998].

@@

The initial discontinuous interpretation has been challenged by subsequent work. For instance, @yoshida2009 investigated whether failure in the Switch task reflects a lack of sound encoding during *familiarization*, or whether it is only due to the nature of the *testing* method which does not allow learning below a certain threshold to be detected. They used the same familiarization procedure as @stager1997, but instead of comparing the looking times in 'same' and 'switch' trials, they tested infants using a two-alternative choice task comparing fixations to target and distractor objects [@golinkoff1987; @fernald2006]. Using this testing method, the researcher found evidence for learning even in 14-month olds.

@@

Some researchers [@swingley2007; @yoshida2009; @pajak2016] proposed that word knowledge may not be encoded in a binary fashion, i.e., it is not the case that children either succeed or fail in encoding minimal contrast when learning the meanings. Rather, they may be encoding this knowledge in a graded fashion (see also @Munakata2001). According to this view, development does not so much involve a qualitative shift (i.e., a sudden emergence of an ability that did not exist before) as much as it consists in the continuous refinement of initially noisy knowledge.

@@

This general framework --- in which the precision of word knowledge is characterized with the variance of a probability distribution --- can already provide an intuitive way of thinking about several findings. In particular, unlike the binary view, the probabilistic view allows for the possibility of word knowledge being both successful and noisy. This new understanding can provide an account for the fact that children show evidence of learning in some testing condition [e.g., @yoshida2009] but not in others [e.g., @stager1997] --- depending on the precision of the measurement.

In a word-pair learning paradigm, children are supposed to associate one label, e.g., "bih", with object 1 and a second label, e.g., "dih", with object 2.  Infants may succeed in learning both associations. Nevertheless, the variance with which the pair of words are encoded can still be large, causing their probability distributions to overlap (Figure \@ref(fig:illus), top). The way this (noisy) knowledge is probed can lead to different results.

In the Switch task [@stager1997], children are understood to succeed if they reject a wrong association (e.g., "bih" with object 2). However, a large overlap between "bin" and "din" means that "bih" is itself a plausible mispronunciation of "dih". The wrong association may not be rejected by children because the speaker could have said "bih" but meant "dih". In the two-alternative choice task [@yoshida2009], children do not have to reject the wrong association; they only need to show a preference for the correct one. Given a label, e.g., "bih", they pick which object is a better match to this label. Though it is possible that the speaker said "bih" and meant "dih", it is *more* likely that the speaker both said and meant "bih" --- this higher probability leads to a preference for the correct object.

In addition to explaining the difference in behavior across the Switch and the preferential looking tasks, the probabilistic account explains difference in behavior within the same task. In particular, when the labels are quite distinct in the perceptual space ("lif" vs. "neem"), the probabilistic distributions do not overlap as much as in the case of similar-sounding words (Figure \@ref(fig:illus), left). This fact means that the learners will have less tolerance for the wrong association, leading to a successful rejection in the Switch task (as was reported by @stager1997 and subsequent studies using the same paradigm). Further, distinctiveness can be enhanced even for minimally different sounds when other cues highlight their difference [@rost2009; @rost2010; @thiessen2007; @yeung09; @dautriche2015].

Finally, development can be understood as an increase in the precision (i.e., a decrease in the variance) of the probabilistic knowledge, leading to a lower overlap between the distributions of similar-sounding words (Figure \@ref(fig:illus), right).  Importantly, a more precise representation still has a non-zero variance --- Learning difficulties can still be induced with challenging stimuli or in cognitively demanding situations as was demonstrated in adults studies [@white2013; @pajak2016].

@@

The present work proposes a model of word-pair learning based on the probabilistic account. We tested the ability of this model to both *explain* various findings in previous experiments in both children and adults (e.g., the fact that similar words are harder to learn than different words) and to *predict* new learning patterns that have not been tested before (i.e., the effect of the referents' similarity on word learning). We explore the extent to which the probabilistic account allows us to understand development in terms of as a continuous refinement in similar representations across the lifespan.

The paper is organized as follows. First, we introduce the model and we explain how it allows us to characterize behavior in a word-pair-learning paradigm. Then we explore the predictions of the model through simulating its behavior across different parameter settings. Next, we quantify the extent to which the model's predictions account for human data we collected from both preschool children and adults. Finally, we discuss the results in the lights of existing accounts of word development.

```{r illus, fig.cap = "An illustration of the probabilistic/continuous account using simulated data. A word is represented with a distribution over the perceptual space (indicated in red or blue). When the uncertainty of the representation is large relative to the distance between the stimuli (top panel), an instance of the red category (indicated with a star) could also be a plausible instance of the green category, hence the low recognition accuracy score. The accuracy is higher when the stimuli are less similar (left panel), or when the representation are more precise (right panel).", fig.align = "center", out.width = "300px"}
knitr::include_graphics("figs/illustration.png")
```

@@

```{r model, fig.cap = "Graphical representation of our model. Circles indicate random variables (shading indicates observed variables). The squares indicate fixed model parameters.", fig.align = "center", out.width = "150px"}
knitr::include_graphics("figs/model.png")
```

@@

\begin{equation} \label{eq:object}
 p(o | C) \sim  \mathcal{N}(\mu_C, \sigma^2_C) 
\end{equation}

\begin{center}
and 
\end{center}

\begin{equation} \label{eq:sound}
p(s| L) \sim  \mathcal{N}(\mu_L, \sigma^2_L)
\end{equation}

Finally, we assume there to be one-to-one mappings between concepts and labels and that observers have successfully learned these mappings during the exposure phase:
\begin{equation}
P(L_i|C_j) = 
\begin{cases}
  1 & \text{if  }  i=j \\  
  0  & \text{otherwise  }
\end{cases}
\end{equation}

@@

In our canonical inference case, the learner hears a sound $s$ and has to decide which object $o$ provides an optimal match to this sound (see Figure \@ref(fig:task)). To this end, they must compute the probability $P(o|s)$ for all possible objects. This probability can be computed by summing over all possible concepts and labels:

\begin{equation}
P(o|s)=\sum_{C,L} P(o, C, L| s)  
\end{equation}

Using the fact that $P(o,C,L|s) = \frac{P(o,C,L,s)}{P(s)}$ and that $P(s)$ does not depend on $o$, we arrive at the equation:

\begin{equation}
P(o|s) \propto \sum_{C,L} P(o, C, L, s) 
\end{equation}

The joint probability $P(o,C,L,s)$ is obtained by factoring the graphical model in Figure \@ref(fig:model):

$$P(o,C,L,s) = P(s|L)P(L|C)P(C|o)P(o)$$

Using Bayes' rule, we can rewrite $P(C|o)$ in terms of $P(o|C)$:

$$P(C|o) = \frac{P(o|C)P(C)}{P(o)}$$

By subtituting this term in the expression of the joint distribution $P(o,C,L,S)$ we obtain:

$$P(o,C,L,s) = P(s|L)P(L|C)P(o|C)P(C)$$

Finally, assuming that the concepts' prior probability $P(C)$ is uniformly distributed,\footnote{This is a reasonable assumption in our particular case given the similarity of the concept pairs used in each naming situation in our experiment.} we obtain the following expression, where all conditional dependencies are now well defined:

\begin{equation} \label{eq:general}
P(o|s) \propto \sum_{C,L}  P(s|L)P(L|C)P(o|C)
\end{equation}

@@

We use the model to predict word learning in a task similar to the one introduced by @stager1997. We used a modified version of the task where the testing method consists in a two-alternative forced-choice [@yoshida2009]. In this task, participants are first exposed to two different word-object pairings (e.g., "lif" - object 1, "neem" - object 2). The word-object associations are introduced sequentially. After this exposure phase, participants perform a series of test trials. In each of these trials, one of the two sounds is uttered (e.g., “lif”) and participants choose the corresponding object from the two alternatives. An overview of the task is shown in Figure \@ref(fig:task).

### Model 1
From the general expression \ref{eq:general}, we derive three exact analytical solutions instantiating different learning assumptions. Recall from expressions \ref{eq:object} and \ref{eq:sound} that $P(o|C)$ and $P(s|L)$ have parameters $\sigma_C$ and $\sigma_L$, respectively, that control perceptual uncertainty. The first solution is derived by assuming that the labels are recovered from sounds with a certain level of uncertainty $\sigma_L > 0$, but that concepts are unambiguously recovered from the observed objects, i.e., $\sigma_C \rightarrow 0$. This assumption has been made --- whether implicitly or explicitly --- by most previous work in this line of research. For example, in @stager1997, the objects were quite dissimilar. Thus, the assumption that they were easily discriminated by infants seems relatively well justified. One important implication of this assumption is that only the similarity of word sounds modulates success in word learning, not the similarity of the referents (as long as these referents are differentiated perceptually). This assumption yields the following probability function:

\begin{equation} \label{eq:model1}
P(o_T|s)= \frac{1}{1 + e^{-\frac{\Delta s^2}{2\sigma_L^2}}}
\end{equation}

### Model 2
The second solution is derived by making the more general assumption that both the labels and the concepts are recovered with noise from the sounds and objects. We first introduce the simplifying assumption that the label-related uncertainty $\sigma_L$ and the concept-related uncertainty $\sigma_C$ are of a similar magnitude, i.e., $\sigma_C \approx \sigma_L = \sigma$. This assumption makes the prediction that the sound similarity and the object similarity impact word learning accuracy in exactly the same way. Furthermore, it allows us to study the behavior of the model with only one free parameter, an important consideration given the small number of datapoints available from any given infant experiment.

\begin{equation} \label{eq:model2}
P(o_T|s)= \frac{1 + e^{- \frac{\Delta s^2 + \Delta o^2}{2\sigma^2}}}{1 + e^{-\frac{\Delta s^2 + \Delta o^2}{2\sigma^2}}+ e^{-\frac{\Delta s^2}{2\sigma^2}} + e^{-\frac{\Delta o^2}{2\sigma^2}}}
\end{equation}

### Model 3
We finally derive the third (and most general) solution which allows label- and concept-related uncertainties to vary independently.

\begin{equation} \label{eq:model3}
P(o_T|s)= \frac{1 + e^{- (\frac{\Delta s^2}{2\sigma_L^2}+ \frac{\Delta o^2}{2\sigma_C^2})}}{1 + e^{-(\frac{\Delta s^2}{2\sigma_L^2}+ \frac{\Delta o^2}{2\sigma_C^2})}+ e^{-\frac{\Delta s^2}{2\sigma_L^2}} + e^{-\frac{\Delta o^2}{2\sigma_C^2}}}
\end{equation}

In order to understand the predictions of the models (especially the more general ones, i.e., Model 2 and 3), Figure \ref{fig:simulation} show simulations of the accuracy $P(o_T|s)$ as a function of the distinctiveness parameters ($\Delta s$ and $\Delta o$) and the uncertainty parameters $\sigma_L$ and $\sigma_C$.

@@

```{r simulation, fig.cap = "The predicted probability of accurate responses in the testing phase as a function of stimuli distinctiveness $\\Delta s$ and $\\Delta o$ and representation precision $\\sigma$ (For simplicity, we use model 2, which assumes that $\\sigma$=$\\sigma_C$=$\\sigma_L$). Dashed line represents chance."}
plot_simulation

@@

2) For fixed values of $\Delta s$ and $\Delta o$, accuracy increases when the representational uncertainty $\sigma$ decreases. This observation provides a simple model for developmental change. Younger children have noisier representations [see @swingley2007; @yoshida2009], which leads to lower word recognition accuracy, especially for similar-sounding words.

@@

### Participants
We report data from $N=$ `r N_good_ch` children ages 4-5 years from the Bing Nursery School on Stanford University’s campus. An additional $N=$ `r N_all_ch-N_good_ch` children participated but were removed from analyses (using preregistered exclusion criteria) because they were not above chance on the catch trials due to the challenging nature of our procedure (see below). We also report data from $N=$ `r N_good_ad` adult participants tested on Amazon Mechanical Turk. An additional $N=$ `r N_all_ad-N_good_ad` were tested but removed from analyses (again, using preregistered exclusion criteria) because they had low scores on the catch trials or because they were familiar with the non-English sound stimuli we used in the adult experiment.

@@

As for the objects, we used the Dynamic Stimuli javascript library\footnote{https://github.com/erindb/stimuli} which allowed us to generate objects in four different categories: "tree," "bird," "bug," and "fish." These categories were described to participants as naturally occurring kinds on an alien planet. In each category, we generated *different*, *intermediate*, and *similar* pairs by manipulating a continuous property controlling features of the category's shape (e.g., body stretch or head fatness).

In order to validate and quantify our similarity scales, we ran a separate survey on Amazon Mechanical Turk where we asked $N=20$ adults participants to evaluate the similarity of each sound and object pair on a 7-point scale. Data are shown in Figure \@ref(fig:stim) where we scaled responses within the range [0,1] for each stimulus group. We used these data in all models as an empirical measurement of the  perceptual distance between the sound pairs and the object pairs. The use of empirical measurement allows us to eliminate $\Delta s$ and $\Delta s$ as free parameters (see @frank2012 and @xu2007 for a similar strategy).

@@

sim_plot <- ggplot(sim_transform, 
      aes(x = labels, y = mean, group = factor(stimuli), col=factor(stimuli))) +
  #geom_point(size=2)+
  geom_line()+
  geom_pointrange(aes(ymin = lower, ymax = upper), 
                 position = position_dodge(width = .1),
                 size=0.2) + 
  scale_colour_discrete(name="Stimuli")+
  theme_few()+
theme(legend.title = element_text(size=7.5),
      legend.text=element_text(size=7.5),
      axis.text = element_text(size = 7.5),
      strip.text = element_text(size = 7.5),
      axis.title = element_text(size = 7.5),
      aspect.ratio = 0.7
      ) +
  ylim(c(-0.1,1.2))+
  scale_y_continuous(breaks = c(0,0.25,0.5, 0.75,1))+
  #ylim(c(0,1.2)) +
  xlab("Stimulus group")+ ylab("Perceptual distance (empirical)")

@@

Each age group saw only two of the three levels of similarity described in the previous sub-section: *different* vs. *intermediate* for the preschoolers, and *intermediate* vs. *similar* for adults. We made this choice in light of pilot studies showing that adults were at ceiling with *different* sounds/objects, and children were at chance with the *similar* sounds/objects. That said, this difference in the level of similarity is accounted for in the model: We used empirical distance measurement to fill in the appropriate values of $\Delta s$ and $\Delta o$ for each age group.

```{r stim, fig.env = "figure", fig.pos = "h", fig.align='center', fig.width=4, fig.height=3, fig.cap = "Distances for both sound and object pairs from an adult norming study. Data represent Likert values normalized to [0,1] interval. Error bars represent 95\\% confidence intervals."}

@@

To maximize our ability to measure subtle stimulus effects, the experiment was a 2x2 within-subjects factorial design with four conditions: high/low sound similarity crossed with high/low visual object similarity. Besides the four conditions, we also tested participants on a fifth catch condition which was similar in its structure to the other ones but was trivially easy and used only to select participants who were able to follow the instructions and show minimal learning.

@@

Preschoolers were tested at the nursery school using a tablet, whereas adults used their own computers to complete the same experiment online. Participants were tested in a random sequence of five conditions: the four experimental conditions plus the catch condition. In each condition, participants saw a first block of four exposure trials followed by four testing trials, and a second block of two exposure trials (for memory refreshment) followed by an additional four testing trials. The length of this procedure was demanding, especially for children, but we adopted a fully within-subjects design based on pilot testing that indicated that precision of measurement was critical for testing our experimental predictions.

In the exposure trials, participants saw two objects associated with their corresponding sounds. We presented the first object on the left side of the tablet's screen simultaneously with the corresponding sound. The second sound-object association followed on the other side of the screen after 500ms. For both objects, visual stimuli were present for the duration of the sound clip (about 800ms).  In the testing trials, participants saw both objects simultaneously and heard only one sound. They completed the trial by selecting which of the two objects corresponded to the sound. The object-sound pairings were randomized across participants, as was the order of the conditions (except for the catch condition which was always placed in the middle of the testing sequence). We also randomized the on-screen position (left vs. right) of the two pictures on each testing trial.\footnote{The experiment can be viewed online at https://tinyurl.com/word-pair-learning-experiment}

```{r allData, fig.env = "figure", fig.pos = "h", fig.align='center', fig.cap = "Accuracy of word recognition as a function of the sound distance, the object distance, and the age group (preschool children vs. adults). We show both the models' predictions (dashed lines) and the experimental results (solid lines, same across the three panels). Error bars represent 95\\% confidence intervals."}

@@

Experimental results are shown in Figure \@ref(fig:allData) (solid lines). We first analyzed the results using a mixed-effects logistic regression with sound distance, object distance and age group as fixed effects, and with a maximal random effects structure (allowing us to take into account the full nested structure of our data) [@barr2013]. We found main effects for all the fixed effects in the regression. For the sound distance, we obtained $\beta =$ `r s_coef` ($p$ `r p_val(s_p)`), replicating previous findings that sound distance modulates success in word learning [e.g., @stager1997].

For object distance, we found $\beta =$ `r c_coef`  ($p$ `r p_val(c_p)`), and this finding confirms the new prediction of our model, according to which, object distance also modulates success in word learning. Note, in particular, that increasing the visual similarity of the objects makes children succeed in learning the similar-sounding words. Finally, for the age group, we obtained  $\beta =$ `r age_coef` ($p$ `r p_val(age_p)`), showing that overall performance improves with age. The full output of the regression model is shown in the Appendix.
